The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2017-03-08 manuscript/manuscript.Rnw
make the core placenta plot two-column

\begin{figure}[tbh]
<<core.placenta.transcriptome,fig.width=9/2.5,fig.height=8,out.width="90mm",out.height="0.8\\textheight",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species) & housekeeping==FALSE,]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]

print(ggplot(cptl[FPKM > 0.1,],aes(x=human_name,y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual("Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=7),
                                            title.position="top"
                                            ),
                         values=0:18)+
      # scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Human Ortholog Name")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      theme(text=element_text(size=9),
            legend.position=c(0.99,0.41),
            axis.text.y=element_text(face="italic",size=7),
            legend.margin=unit(0,"mm"),
            legend.justification=c(1,1),
            plot.margin=unit(c(0,0,0,0),"mm")
            ))
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  ortholog pairs of \human genes expressed (\ac{FPKM}$ \ge 10$) in the
  placenta of all 13 eutherians studied with housekeeping genes
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited} removed
  ordered (from top to bottom) by median expression. If we also require
  expression (\ac{FPKM}$ \ge 10$) in the marsupial, \opossum, there are 
  \Sexpr{cpt.distribution.shape[`Expression (FPKM)` > 10 &
    `Min. Exp. in` == '14 of 14 Species',max(num_genes)]} genes
  which meet this criteria.
  \Cref{fig:core_placenta_transcriptome_shape} shows the shape of the
  distribution of the number of genes in the core placenta
  transcriptome if different \ac{FPKM} and species restriction criteria are
  utilized.}
\label{fig:core_placenta_transcriptome}
\end{figure}

>>>>

\begin{figure}[tbh]
<<core.placenta.transcriptome,fig.width=19/2.5,fig.height=6,out.width="190mm",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species) & housekeeping==FALSE,]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]
cptl <- cptl[,median_FPKM:=median(FPKM),by=human_name]
cptl <- cptl[,facet:="high"]
cptl <- cptl[median_FPKM < median(median_FPKM),facet:="low"]
print(ggplot(cptl[FPKM > 0.1,],aes(x=human_name,y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(breaks=c(10^1,10^2,10^3,10^4),
                    labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual(NULL,##"Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=7),
                                            title.position="top"
                                            ),
                         values=0:18)+
      scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Human Ortholog Name")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      expand_limits(y=c(10^4)) + 
      theme(text=element_text(size=7.5),
            legend.position=c(0.99,0.01),
            axis.text.y=element_text(face="italic",size=7.5),
            legend.margin=margin(0,0,0,0,unit="mm"),
            legend.spacing=unit(0,"mm"),
            legend.justification=c(1,0),
            plot.margin=unit(c(0,0,0,0),"mm"),
            strip.background = element_blank(),
            strip.text.x = element_blank()
            )
      + facet_wrap(~facet,scales="free")
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
  ortholog pairs of \human genes expressed (\ac{FPKM}$ \ge 10$) in the
  placenta of all 13 eutherians studied with housekeeping genes
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited} removed
  ordered (from top to bottom) by median expression. If we also require
  expression (\ac{FPKM}$ \ge 10$) in the marsupial, \opossum, there are 
  \Sexpr{cpt.distribution.shape[`Expression (FPKM)` > 10 &
    `Min. Exp. in` == '14 of 14 Species',max(num_genes)]} genes
  which meet this criteria.
  \Cref{fig:core_placenta_transcriptome_shape} shows the shape of the
  distribution of the number of genes in the core placenta
  transcriptome if different \ac{FPKM} and species restriction criteria are
  utilized.}
\label{fig:core_placenta_transcriptome}
\end{figure}
